---
title: "Computational Musicology Portfolio"
author: "Gijs de Huu"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
editor: visual
execute:
  echo: false
  warning: false
  message: false
---
<style>
body {
  background-color: #f6f7fb;
}
</style>

# Background and Corpus

This portfolio presents a computational comparison between two electronic dance genres: Italo disco (1979–1986) and Berlin minimal techno (2000–2024). The goal is to investigate how these genres differ not only in track-level Spotify features, but also in deeper structural properties such as harmony, timbre, loudness, and rhythmic organisation.

The corpus consists of two curated playlists of 22 tracks each. Although these playlists were not intended to include every important track in each genre, the portfolio assumes that the selected tracks are sufficiently representative of the broader styles. So any conclusions should be seen first as conclusions about this specific set of songs, and only more carefully as conclusions about the genres as a whole.

Two representative tracks were selected for closer analysis: *Two of Hearts* by Stacey Q and *Levo* by Recondite. I selected these tracks because they featured the musical differences I wanted to examine computationally. Although *Two of Hearts* is an American rather than Italian track, it was included because it clearly shows key Italo Disco features, including looping minor-key harmonies, synthesizer-driven textures, and repetitive chord-based structure. Additionally, the genre itself was not limited to Italy alone, since it developed into an international phenomenon in the early 1980s. *Levo* was chosen as a representative example of Berlin Minimal Techno because of its emphasis on sustained tonal centres, gradual development, and timbral transformation.

The portfolio combines Spotify audio features from Exportify, including danceability, loudness, valence, and tempo, with time-based features such as chroma, MFCCs, and tempograms extracted in Sonic Visualiser and analysed in R Studio using compmus. Spotify audio features are useful but imperfect descriptors. They make corpus-level comparison possible, but they do not capture every musically meaningful aspect of genre or listening experience.

The central question of this portfolio is whether the difference between Italo Disco and Berlin Minimal Techno can be captured computationally, and if so, whether this lies in surface features or in more underlying patterns of pitch, timbre, and repetition over the course of the track.

# Track-Level Features

## Tempo and Danceability

This first comparison focuses on tempo (BPM) and danceability. Rather than assuming that tempo directly determines danceability, this analysis asks how these two features are distributed across the two genres.

```{r}
library(tidyverse)

italo <- read_csv("ITAD.csv") %>%
  mutate(genre = "Italo Disco")

minimal <- read_csv("BMT.csv") %>%
  mutate(genre = "Berlin Minimal Techno")

corpus <- bind_rows(italo, minimal)
genre_counts <- corpus %>%
  count(genre)

genre_correlations <- corpus %>%
  group_by(genre) %>%
  summarise(
    n_tracks = n(),
    correlation = cor(Tempo, Danceability, use = "complete.obs")
  )

ggplot(corpus, aes(x = Tempo, y = Danceability)) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~ genre) +
  labs(
    title = "Tempo and Danceability by Genre",
    subtitle = "Track-level comparison",
    x = "Tempo (BPM)",
    y = "Danceability (0–1)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    strip.background = element_rect(fill = "white", colour = NA),
    strip.text = element_text(face = "bold")
  )
```

The plot shows that both playlists fall within tempo ranges typical of dance music and include many tracks with relatively high danceability values. The clearest difference is that Berlin Minimal Techno is concentrated within a narrower tempo range, whereas Italo Disco shows more variation in tempo. This suggests that the figure is more informative about differences in tempo spread than about any strong relationship between tempo and danceability itself.

Because danceability is shaped by more than tempo alone, the following sections move beyond track-level features to examine pitch organisation, timbre, loudness, and rhythm in more detail.

# Chroma / Pitch

:::::: column-page
To compare harmonic structure at the level of local pitch content, chromagrams were generated for one representative track from each playlist: *Two of Hearts* (Italo Disco) and *Levo* (Berlin Minimal Techno). Chroma features show the intensity of each pitch class (C, C#, D, etc.) over time, independent of octave. They were extracted in Sonic Visualiser and exported as CSV files.

```{r}
library(tidyverse)
library(compmus)

pitch_order <- c("C","C#","D","Eb","E","F","F#","G","Ab","A","Bb","B")

two <- read_csv("two-chroma.csv", show_col_types = FALSE)
levo <- read_csv("levo-chroma.csv", show_col_types = FALSE)

colnames(two)[1] <- "Time"
colnames(levo)[1] <- "Time"

two_long <- two %>%
  pivot_longer(cols = -Time, names_to = "Pitch", values_to = "Magnitude") %>%
  mutate(Pitch = factor(Pitch, levels = pitch_order)) %>%
  drop_na(Time)

levo_long <- levo %>%
  pivot_longer(cols = -Time, names_to = "Pitch", values_to = "Magnitude") %>%
  mutate(Pitch = factor(Pitch, levels = pitch_order)) %>%
  drop_na(Time)
```

## Chromagrams of the Two Tracks

::::: columns
::: {.column width="50%"}
```{r}
ggplot(two_long, aes(x = Time, y = Pitch, fill = Magnitude)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    title = "Two of Hearts – Stacey Q",
    x = "Time (s)",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA)
  )
```
:::

::: {.column width="50%"}
```{r}
ggplot(levo_long, aes(x = Time, y = Pitch, fill = Magnitude)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    title = "Levo – Recondite",
    x = "Time (s)",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA)
  )
```
:::
:::::

The chromagrams already show a clear difference in harmonic organisation. *Two of Hearts* shows multiple pitch classes being active at the same time, forming recurring vertical groupings that suggest chord-based harmony and repeated harmonic cycles. This fits the style of Italo Disco, where stable chord progressions and tonal repetition often play a central role.

On the other hand, *Levo* shows fewer simultaneously active pitch classes, each of them often sustained over a longer time span. This creates a more horizontally continuous pattern, consistent with harmonic stasis, the prolonged use of a limited set of pitches with gradual variation over time.

## Chroma Self-Similarity

While the chromagrams show which pitch classes are active at each moment, a chroma self-similarity matrix (SSM) shows how similar different moments in the same track are to one another. This makes it easier to identify repeated sections and long-distance harmonic recurrences.

```{r chroma-ssm-combined, fig.width=13, fig.height=6}
make_chroma_ssm <- function(path, label, downsample = 10L,
                            norm = "manhattan", dist = "cosine") {

  df <- read_csv(path, show_col_types = FALSE)
  colnames(df)[1] <- "TIME"

  chroma_cols <- c("C","C#","D","Eb","E","F","F#","G","Ab","A","Bb","B")

  dt <- median(diff(df$TIME), na.rm = TRUE)

  df %>%
    mutate(
      start = TIME,
      duration = lead(TIME) - TIME,
      duration = if_else(is.na(duration), dt, duration),
      chroma = pmap(select(., all_of(chroma_cols)), c)
    ) %>%
    select(start, duration, chroma) %>%
    filter(row_number() %% downsample == 0L) %>%
    mutate(chroma = map(chroma, ~ compmus_normalise(.x, norm))) %>%
    compmus_self_similarity(chroma, dist) %>%
    mutate(track = label)
}

two_chroma_ssm  <- make_chroma_ssm("two-chroma.csv",  "Two of Hearts", downsample = 10L)
levo_chroma_ssm <- make_chroma_ssm("levo-chroma.csv", "Levo",          downsample = 10L)

chroma_both <- bind_rows(two_chroma_ssm, levo_chroma_ssm)

ggplot(chroma_both, aes(
  x = xstart + xduration / 2,
  y = ystart + yduration / 2,
  fill = d
)) +
  geom_tile() +
  facet_wrap(~ track, nrow = 1) +
  coord_equal(expand = FALSE) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Chroma Self-Similarity",
    subtitle = "Pitch-class similarity over time",
    x = "Time",
    y = "Time"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    strip.background = element_rect(fill = "white", colour = NA),
    strip.text = element_text(face = "bold")
  )
```

The chroma SSMs make the structural difference easier to see. In *Two of Hearts*, repeated off-diagonal patterns show that similar chord patterns return in different parts of the song. This suggests that harmonic repetition is an important part of the track’s structure.

In *Levo*, the pattern of long-distance harmonic recurrence is weaker and more spread out. This does not mean the track lacks structure, but rather that its form depends less on changing chord progressions and more on repetition within a limited tonal space. In other words, the track is harmonically restrained even when it remains strongly patterned.

Taken together, the chromagrams and chroma SSMs suggest that *Two of Hearts* is more strongly organised by repeating harmonic cycles, while *Levo* relies more on sustained tonal centres and gradual change.

## Note on Key Estimation

A further key-estimation analysis was also explored using pitch-template matching, following the course material on chordograms and keygrams. Although this provided some support for the contrast between tonal stability in *Two of Hearts* and greater ambiguity in *Levo*, the resulting visualisation was less interpretable than the chromagram and chroma self-similarity matrix. For that reason, the pitch-based comparison in this portfolio focuses on those clearer representations.
::::::

# Timbre

## Timbre Self-Similarity (MFCC)

As an addition to the pitch-based comparison, I also computed self-similarity matrices based on MFCC features. Whereas chroma focuses on pitch-class content, MFCCs capture timbral properties such as spectral shape and sound colour. This makes them useful for showing whether sections are connected through similar texture and production qualities, even when harmonic content remains relatively stable.

```{r timbre-ssm-combined, message=FALSE, warning=FALSE}
make_timbre_ssm <- function(path, label, downsample = 50L,
                            norm = "euclidean", dist = "cosine") {
  mfcc <- read_csv(path, show_col_types = FALSE)
  colnames(mfcc)[1] <- "TIME"

  dt <- median(diff(mfcc$TIME), na.rm = TRUE)

  mfcc %>%
    mutate(
      start = TIME,
      duration = lead(TIME) - TIME,
      duration = if_else(is.na(duration), dt, duration),
      timbre = pmap(select(., starts_with("BIN")), c)
    ) %>%
    select(start, duration, timbre) %>%
    filter(row_number() %% downsample == 0L) %>%
    mutate(timbre = map(timbre, ~ compmus_normalise(.x, norm))) %>%
    compmus_self_similarity(timbre, dist) %>%
    mutate(track = label)
}

two_timbre_ssm  <- make_timbre_ssm("two-mfcc.csv",  "Two of Hearts", downsample = 50L)
levo_timbre_ssm <- make_timbre_ssm("levo-mfcc.csv", "Levo",          downsample = 50L)

timbre_both <- bind_rows(two_timbre_ssm, levo_timbre_ssm)

ggplot(timbre_both, aes(
  x = xstart + xduration / 2,
  y = ystart + yduration / 2,
  fill = d
)) +
  geom_tile() +
  facet_wrap(~ track, nrow = 1) +
  coord_equal(expand = FALSE) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Timbre Self-Similarity (MFCC)",
    subtitle = "Similarity of sound colour and texture over time",
    x = "Time",
    y = "Time"
  ) +
 theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    strip.background = element_rect(fill = "white", colour = NA),
    strip.text = element_text(face = "bold")
  )
```

In *Levo*, the block-like regions and clearer boundaries suggest more noticeable timbral change across sections, which is consistent with Berlin Minimal Techno.

In *Two of Hearts*, the timbral SSM appears more consistent across time, suggesting a more stable overall sound between sections.

Together, the pitch- and timbre-based SSMs show that the two tracks organise repetition differently. *Two of Hearts* relies more on harmonic repetition, while *Levo* relies more on timbral change.

# Loudness

## Loudness and Energy

Loudness (measured in decibels) offers another way to understand how musical intensity is spread across tracks. Loudness reflects the overall energy level of a track and can contribute to whether the music feels more varied or more steady.

```{r}
library(tidyverse)

italo <- read_csv("ITAD.csv", show_col_types = FALSE) %>%
  mutate(genre = "Italo Disco")

minimal <- read_csv("BMT.csv", show_col_types = FALSE) %>%
  mutate(genre = "Berlin Minimal Techno")

corpus <- bind_rows(italo, minimal)

ggplot(corpus, aes(x = genre, y = Loudness, fill = genre)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Loudness Distribution by Genre",
    x = "",
    y = "Loudness (dB)"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```

This boxplot compares loudness across the two genres. Both show a broadly similar range, which is expected for electronically produced dance music. Any difference here appears to be one of distribution rather than a sharp contrast between genres. Loudness therefore complements the broader comparison, but on its own it does not distinguish the playlists as clearly as the pitch-, timbre-, or rhythm-based analyses.

# Rhythm / Tempo

## Rhythmic Structure: Tempograms

To analyse rhythmic structure, tempograms were computed for both tracks. A tempogram shows how strongly different tempo periodicities (in BPM) are present over time. Brighter regions indicate stronger rhythmic regularity at a given tempo.

```{r}
library(tidyverse)

plot_tempogram <- function(file, label) {

  read_csv(file, show_col_types = FALSE) |>
    rename(TIME = 1) |>
    pivot_longer(-TIME, names_to = "tempo", values_to = "value") |>
    mutate(tempo = as.numeric(tempo)) |>
    ggplot(aes(x = TIME, y = tempo, fill = value)) +
    geom_tile() +
    scale_fill_viridis_c() +
    labs(
      title = paste("Tempogram:", label),
      x = "Time (s)",
      y = "Tempo (BPM)"
    ) +
    coord_cartesian(ylim = c(40, 300)) +
    theme_minimal(base_size = 14) +
    theme(
      panel.grid = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA)
    )
}

plot_tempogram("twohearts-tempogram.csv", "Two of Hearts")
plot_tempogram("levo-tempogram.csv", "Levo")
```

The tempograms show how rhythmic periodicities are distributed over time. In both tracks, a strong grouping of energy appears around the 120–130 BPM range, visible as a horizontal band of greater intensity. This confirms that both tracks operate within a dance-oriented tempo range.

The difference lies in stability. In *Levo*, the band around the main tempo is more continuous and regular over time, suggesting a more persistent rhythmic structure. In *Two of Hearts*, the same tempo range is present but less uniform, which suggests greater rhythmic variation. This supports the broader contrast developed throughout the portfolio: although both tracks are danceable and tempo-relatedly similar, *Levo* depends more on rhythmic persistence, whereas *Two of Hearts* allows more local variation within that framework.

# Clustering or Classification

## Classification of Italo Disco and Berlin Minimal Techno

As a final step, I trained a classifier on Spotify track-level audio features to test whether Italo Disco and Berlin Minimal Techno can be distinguished computationally. This extends the earlier visual analyses by asking whether the differences in pitch, timbre, and rhythm also appear in higher-level track features.

To do this, a random forest classifier was trained on the two playlists. I first used a broader set of Spotify features, and then a smaller feature set to see whether similar performance could be achieved with fewer variables.

```{r classification-setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(ranger)

set.seed(123)

italo <- read_csv("ITAD.csv", show_col_types = FALSE) %>%
  mutate(genre = "Italo")

techno <- read_csv("BMT.csv", show_col_types = FALSE) %>%
  mutate(genre = "Techno")

tracks <- bind_rows(italo, techno) %>%
  mutate(genre = factor(genre)) %>%
  drop_na(
    Danceability, Energy, Loudness, Speechiness, Acousticness,
    Instrumentalness, Liveness, Valence, Tempo, `Duration (ms)`
  )

tracks %>% count(genre)
```

### Full-feature classifier

```{r full-model, warning=FALSE, message=FALSE}
split <- initial_split(tracks, prop = 0.75, strata = genre)
train_data <- training(split)
test_data  <- testing(split)

genre_recipe <- recipe(
  genre ~ Danceability + Energy + Loudness + Speechiness +
    Acousticness + Instrumentalness + Liveness + Valence +
    Tempo + `Duration (ms)`,
  data = train_data
) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())

rf_model <- rand_forest(
  trees = 500,
  mtry = 3,
  min_n = 2
) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")

rf_workflow <- workflow() %>%
  add_recipe(genre_recipe) %>%
  add_model(rf_model)

rf_fit <- fit(rf_workflow, data = train_data)

rf_preds <- predict(rf_fit, test_data) %>%
  bind_cols(test_data %>% select(genre))

conf_mat_full <- conf_mat(rf_preds, truth = genre, estimate = .pred_class)
acc_full <- accuracy(rf_preds, truth = genre, estimate = .pred_class)

conf_mat_full
acc_full
```

### Reduced-feature classifier

```{r reduced-model, warning=FALSE, message=FALSE}
genre_recipe_small <- recipe(
  genre ~ Danceability + Energy + Loudness + Instrumentalness +
    Valence + Tempo,
  data = train_data
) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())

rf_workflow_small <- workflow() %>%
  add_recipe(genre_recipe_small) %>%
  add_model(rf_model)

rf_fit_small <- fit(rf_workflow_small, data = train_data)

rf_preds_small <- predict(rf_fit_small, test_data) %>%
  bind_cols(test_data %>% select(genre))

conf_mat_small <- conf_mat(rf_preds_small, truth = genre, estimate = .pred_class)
acc_small <- accuracy(rf_preds_small, truth = genre, estimate = .pred_class)

conf_mat_small
acc_small
```

### Feature importance

```{r feature-importance, warning=FALSE, message=FALSE}
importance_df <- extract_fit_parsnip(rf_fit)$fit$variable.importance %>%
  enframe(name = "feature", value = "importance") %>%
  arrange(desc(importance))

ggplot(importance_df, aes(x = reorder(feature, importance), y = importance)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Feature Importance in the Full Random Forest Model",
    x = NULL,
    y = "Importance"
  ) +
  theme_minimal(base_size = 14)
```

The full-feature classifier tests whether the two playlists can be separated using Spotify audio features alone. The confusion matrix shows how many test tracks were correctly or incorrectly assigned to each genre, while the accuracy score summarises overall performance.

The reduced-feature model asks whether similar results can be achieved with a smaller set of variables. In this sample, the reduced model performs just as well as the full model, which suggests that a smaller group of features already captures much of the difference between the two playlists.

The feature-importance plot indicates which variables contributed most strongly to the full model. Instrumentalness, valence, loudness, danceability, and tempo appear especially important. This suggests that some of the contrast between the playlists is reflected not only in detailed structural analyses, but also in broader Spotify feature profiles.

Because the dataset contains only 22 tracks per playlist, these results should be interpreted cautiously. The classifier’s performance is informative for this sample, but it should not be treated as definitive evidence that the genres are always cleanly separable.

## Hierarchical Clustering

In addition to classification, I used hierarchical clustering to examine whether tracks from the two playlists also group together without genre labels. This addresses the same question in a different way: instead of giving the model genre labels, it checks whether Italo Disco and Berlin Minimal Techno naturally form distinct clusters based on Spotify audio features alone.

```{r clustering-setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(stringr)
library(tibble)
library(ggdendro)

italo <- read_csv("ITAD.csv", show_col_types = FALSE) %>%
  mutate(genre = "Italo")

techno <- read_csv("BMT.csv", show_col_types = FALSE) %>%
  mutate(genre = "Techno")

tracks <- bind_rows(italo, techno) %>%
  mutate(
    label = paste(str_trunc(`Track Name`, 28), genre, sep = " | ")
  ) %>%
  drop_na(
    Danceability, Energy, Loudness, Speechiness, Acousticness,
    Instrumentalness, Liveness, Valence, Tempo, `Duration (ms)`
  )

tracks %>% count(genre)
```

### Pre-processing

```{r clustering-prep, warning=FALSE, message=FALSE}
cluster_data <-
  recipe(
    label ~ Danceability + Energy + Loudness + Speechiness +
      Acousticness + Instrumentalness + Liveness + Valence +
      Tempo + `Duration (ms)`,
    data = tracks
  ) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  prep() %>%
  juice() %>%
  column_to_rownames("label")
```

### Distance matrix

```{r clustering-distance, warning=FALSE, message=FALSE}
track_dist <- dist(cluster_data, method = "euclidean")
```

### Dendrograms with different linkage methods

```{r clustering-dendrograms, warning=FALSE, message=FALSE, fig.height=10}
par(mfrow = c(3, 1), mar = c(3, 4, 2, 1))

plot(
  hclust(track_dist, method = "single"),
  main = "Hierarchical Clustering - Single Linkage",
  xlab = "",
  sub = "",
  cex = 0.6
)

plot(
  hclust(track_dist, method = "average"),
  main = "Hierarchical Clustering - Average Linkage",
  xlab = "",
  sub = "",
  cex = 0.6
)

plot(
  hclust(track_dist, method = "complete"),
  main = "Hierarchical Clustering - Complete Linkage",
  xlab = "",
  sub = "",
  cex = 0.6
)

par(mfrow = c(1, 1))
```

### Final dendrogram

```{r clustering-final, warning=FALSE, message=FALSE, fig.height=7}
hc_complete <- hclust(track_dist, method = "complete")

ggdendrogram(hc_complete, rotate = TRUE, size = 2) +
  labs(
    title = "Hierarchical Clustering of Italo Disco and Berlin Minimal Techno",
    subtitle = "Complete linkage on scaled Spotify audio features",
    x = "",
    y = "Height"
  ) +
  theme_minimal(base_size = 10)
```

The dendrogram shows how tracks group together based on their overall similarity across Spotify audio features. Rather than using genre labels directly, clustering asks whether the tracks form groups on their own.

In this case, the dendrogram suggests partial separation rather than a clean split. Some tracks cluster with others from the same playlist, but Italo Disco and Berlin Minimal Techno do not form two fully distinct branches. This indicates that the Spotify feature profiles capture some difference between the playlists, while also showing overlap between individual tracks.

This makes clustering a useful complement to classification. Whereas the classifier tests whether a model can learn to distinguish the two playlists, the dendrogram shows that the underlying feature space is structured by genre only to a limited extent, not in an absolute way.

# Reflection

This portfolio showed that Italo Disco and Berlin Minimal Techno differ not only in broad Spotify features, but also in how they organise pitch, timbre, rhythm, and repetition over time. Beginning with track-level Spotify features and moving towards more detailed representations such as chromagrams, self-similarity matrices, and tempograms, the aim was to understand how these genres differ in harmony, timbre, rhythm, and overall structure.

Across the analyses, a consistent contrast emerged. *Two of Hearts* is organised more strongly through harmonic repetition, with clear chord cycles and recurring tonal patterns. This was visible in the chromagram and reinforced by the chroma self-similarity matrix. By contrast, *Levo* relies less on harmonic progression and more on gradual development within a limited tonal space.

This contrast becomes even clearer in timbre. The MFCC-based self-similarity matrices showed that *Levo* is structured through changes in texture and sound colour, with more clearly differentiated timbral sections. *Two of Hearts*, on the other hand, maintains a more consistent timbral identity, which suggests that its form depends less on timbral contrast and more on harmonic repetition.

The rhythmic analyses showed that both tracks fall within a similar tempo range typical of dance music, but differ in how rhythm functions structurally. The loudness analysis supported this broader contrast by showing differences in how musical intensity was distributed across the two playlists.

Finally, the classification and clustering analyses showed that these differences are not only visible in detailed structural representations, but also in broader Spotify audio features. Although the dataset is relatively small, the fact that the genres can be distinguished computationally suggests that the differences described throughout the portfolio are based on consistent patterns rather than chance.

A main contribution of this portfolio is that it brings together several computational approaches into one stylistic comparison. By combining track-level features with pitch, timbre, rhythm, and grouping methods, it shows how measurable musical patterns can be connected to differences in listening experience between genres.

## Acknowledgements

I would like to thank my fellow students for their feedback throughout the course, which helped improve both the clarity and interpretation of this portfolio. I would also like to thank Atser Damsma for his guidance and feedback during the course.
